### 2024-02-15 outputs/paper/paper.qmd
added estimand paragraph in intro , deleted files

---
title: "Political Polarization, Public Trust, and Healthcare Disparities: Unraveling COVID-19's Impact and Future Preparedness"
author: 
  - Adrian Ly
  - Hannah Yu
  - Sakhil Goel
thanks: "Code and data are available at: https://github.com/hannahyu07/US-Covid-Analysis.git 
This reproduction was performed after a replication on the Social Science Reproduction platform: [link here](https://www.socialsciencereproduction.org/reproductions/c35e8e98-762d-4c00-a1a3-544ed0b2008d/index)"
date: today
date-format: long
abstract: "The COVID-19 pandemic has exposed deep disparities in health outcomes across various regions in the United States. Using research conducted by @citearticle, we analyze how variances in local health infrastructure and socio-political factors influenced COVID-19 outcomes. Our analysis reveals that political polarization notably impacted adherence to health guidelines, particularly in Republican-leaning states, correlating with higher death rates. Furthermore, disparities in healthcare access and housing significantly affected the life expectancy of different ethnic groups. These findings emphasize the need for tailored public health strategies to address infrastructure and socio-political deficiencies, while future research should continue to explore these factors for better pandemic preparedness and response."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

#### Workspace setup ####
library(tidyverse)
library(sf)
library(readxl)
library(knitr)
library(janitor)
library(lubridate)
library(dplyr)
library(data.table)
library(RColorBrewer)
library(ggpubr)
library(scales)
library(here)
library(kableExtra)
library(gt)

#### Reading files ####

ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
vote_share <- read.csv(here::here("inputs/data/Vote_share_2020_data.csv"), stringsAsFactors = FALSE)
deaths <- read.csv(here::here("inputs/data/covid-19-death-rates.csv"), stringsAsFactors = FALSE)

```

# Introduction

The COVID-19 pandemic has exposed significant disparities in health outcomes across various regions of the United States, drawing attention to the intricate interplay of local factors in shaping disease trajectories. Understanding these dynamics is crucial for informing effective public health interventions and fostering resilience in the face of future pandemics. In this paper, building upon prior research, we delve into the multifaceted relationship between local health infrastructure, socio-political factors, and COVID-19 outcomes, aiming to provide insights into the mechanisms driving disparities and identify strategies for targeted intervention.

Our analysis seeks to estimate several key parameters, including the impact of political polarization on adherence to health guidelines, the association between social vulnerabilities (such as healthcare access, housing conditions, and ethnic demographics) and disparities in COVID-19 outcomes, and the influence of community health systems on disease trajectories. These estimands guide our methodology, data analysis, and interpretation of results, allowing us to comprehensively address the research questions.

Our analysis is grounded in comprehensive data sourced from the paper, **"Why Did the Best Prepared Country in the World Fare So Poorly during COVID?"** [@citearticle], enabling a nuanced exploration of how community health systems and socio-political contexts intersect to influence disease outcomes. While leveraging the initial findings from @citearticle, we expand the scope of inquiry to investigate broader implications and unearth deeper insights into the complexities of COVID-19 disparities. We particularly focus on the impact of political polarization on adherence to health guidelines and the role of social vulnerabilities such as healthcare access and housing conditions on the life expectancy of various racial and ethnic groups. One key finding that emerges from our study is the influence of political polarization, with Republican-leaning states exhibiting slightly higher clustering around elevated death rates, highlighting the importance of socio-political factors in shaping public health responses.

[@fig-GHS] gives us some context to motivate our research. This map allows us to visualize the Global Health Security (GHS) Index which is designed to assess a country's capabilities to handle significant infectious disease outbreaks. A higher on the GHS Index indicates a stronger health system capacity for pandemic readiness. The GHS Index evaluates countries' readiness for pandemics and other significant biological threat emergencies, using a comprehensive framework. This framework includes measurements grouped into six broad categories: prevention, detection, response, health system capacity, compliance with international norms, and risk environment. Higher scores suggest that a country has a more robust health system capacity, with better capabilities to prevent, detect, respond to, and mitigate the spread of an epidemic.

Conversely, a lower GHS Index score implies weaker health system capacities, suggesting that the country may be less prepared to handle pandemics and other significant biological threats. Lower scores indicate deficiencies in one or more of the critical areas assessed by the Index, which could potentially hinder the country's ability to effectively manage and control the spread of infectious diseases.

As we navigate the complexities of the COVID-19 pandemic, it is imperative to not only understand the factors contributing to disparities but also to identify actionable strategies for mitigating these inequities and strengthening community resilience. In the subsequent sections of this paper, we outline our research process, findings, and implications as follows: @sec-data delves into our methodology, encompassing data collection, cleaning, and measurement techniques. In @sec-results, we present detailed findings accompanied by graphs, tables, and other visual aids for clarity. Following this, @sec-discussion engages in a discussion on the significance and implications of our results for public health policy and practice, addressing unexpected findings, limitations, and avenues for further research. Finally, @sec-conclusion offers a conclusion and recommendations, summarizing key insights and providing actionable recommendations for policymakers, practitioners, and future research directions. By elucidating these critical relationships and expanding upon prior research, our study contributes to the broader discourse on pandemic preparedness and underscores the importance of tailored interventions in addressing underlying structural determinants of health disparities.

```{r}
#| echo: false
#| eval: true
#| label: fig-GHS
#| fig-cap: Global Health Security Index Scores by Country
#| warning: false

ghs_data <- ghs_data %>%
  mutate(COUNTRY = case_when(
    Country == "United States of America" ~ "United States",
    Country == "Bolivia (Plurinational State of)" ~ "Bolivia",
    # Add other necessary name changes here
    TRUE ~ Country
  ))

world_ghs <- merge(world_map, ghs_data, by.x = "COUNTRY", by.y = "COUNTRY", all.x = TRUE)

ggplot(data = world_ghs) +
  geom_sf(aes(fill = `OVERALL SCORE`), color = "white") +
  scale_fill_distiller(palette = "Spectral", na.value = "lightgrey", 
                       name = "GHS Index Score", direction = 1,
                       limits = c(min(ghs_data$`OVERALL SCORE`, na.rm = TRUE), max(ghs_data$`OVERALL SCORE`, na.rm = TRUE)),
                       labels = comma) +  # Use comma to format numbers with commas
  theme_minimal() +
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5))

```

# Data {#sec-data}

## Source

The datasets utilized in the `original paper` [@citearticle] and this paper were mainly compiled from various sources such as the @data1, @data2, @data3, @data4, @data5, @data6, @data7, @data8, @data9, @data10, and @data11. Additionally, to address the original paper's lack of US COVID statistics and political party support data, we incorporated information from @citedata1 and @citedata2. The political party support data we utilized from @citedata1 was compiled from @data12.

@citedata1 discusses the long-term impacts of COVID-related school closures. From this source, we utilized the dataset on voting shares during the 2020 election by county.
@citedata2 summarizes COVID-19 death rates in the United States as of March 2023, organized by state. Analyzing results from both datasets allows us to explore the relationship between political affiliation and COVID-19 outcomes. Our reproduction aims to fill these gaps and includes tables and graphs not presented in the original paper to support our findings.

## Methodology

This paper aims to replicate the data originally presented by @citearticle, as mentioned previously. Subsequently, we will integrate this replication into our own data analysis, incorporating COVID statistics and political party support data. Our objective is to leverage both the initial replication and our subsequent analysis to contribute a fresh perspective to the original paper, enriching it with additional data. In contrast to the original paper, which extensively discussed the inadequacies of US preparedness compared to other countries, we specifically diverge our focus to examine the political and socioeconomic factors that influenced the US's COVID outcomes.

`R` [@citeR] was the language and environment used for the bulk of this analysis, alongside `tidyverse` [@citeTidyVerse], `sf` [@citeSF], `readxl` [@citeReadXL], `knitr` [@citeKnitR], `janitor` [@citeJanitor], `lubridate` [@citeLubridate], `dplyr` [@citeDplyr], `data.table` [@citeDataTable], `RColorBrewer` [@citeRColorBrewer], `ggpubr` [@citeggpubr], `ggplot2` [@citeggplot], `here` [@citeHere], `kableExtra` [@citeKableExtra], `webshot` [@citewebshot], `webshot2` [@citewebshot2], `gt` [@citegt] and `scales` [@citeScales].

## Data cleaning

From all the data provided from @citearticle, @citedata1, and @citedata2, we mainly utilize  @data3, @data8, @data12, and @citedata2. The data from @citearticle is well-organized and require minimal cleaning for analysis.

The data collected from @data12 is at the county level as depicted on @tbl-sample. However, since our analysis focuses on people's political preferences at the state level, we create a new data frame called "vote_data_by_state" to aggregate voting data by state. We utilize the "group_by" function to generate new variables at the state level, calculating the total votes, total Republican party votes, total Democratic party votes, and the total vote difference between the two parties by county. Additionally, We also create variables for the average percentage of votes each party receives and the difference by each state.

To determine the total number of votes in the Top 10 states with the highest proportion of Republican votes, we create another new data frame called "sorted_states" from the "vote_data_by_state" data frame. We include a new variable "proportion_republican" that calculates the proportion of votes the Republican party receives from each state and orders the states by the highest proportion of Republican votes.

```{r}
#| echo: false
#| eval: true
#| label: tbl-sample
#| tbl-cap: Sample of Vote Share Data
#| warning: false

# Take a sample of the data (let's say the first 5 rows)
vote_share_sample <- head(vote_share, 5)

# Print the sample table
kable(vote_share_sample, 
      format = "markdown", 
      col.names = c("State Name", "County  FIPS", "County Name", "Votes GOP", "Votes DEM", "Total Votes", "Differ- ence", "Percent- age GOP", "Percent- age DEM", "Percentage Point Difference")
)

```

The data from @data2 lacks column names for each column. To retrieve data by column, we assign column names "State" and "Death_Rate" to the variables of interest. Subsequently, to find the top 10 states with the highest death rates, we rank the states based on death rate. Lastly, to create the final figure that incorporates the COVID death rate and people's political preferences by state, we merge the two datasets from @data12 and @citedata2 and proceed with our analysis.

## Data Measurement

The transformation of real-world phenomena into quantifiable data points in our dataset is influenced by various factors, leading to inherent limitations in our  measurements for COVID death rates, life expectancy rates [@citedefenition], and political affiliation.

The determination of whether someone is deceased and whether they had COVID involves several steps and considerations. Medical professionals typically confirm a person's death through various means, including clinical examination, vital signs assessment, and sometimes laboratory tests. For COVID-related deaths, individuals are often tested for the presence of the virus through diagnostic tests such as PCR or antigen tests. Additionally, medical history and symptoms consistent with COVID-19 may also contribute to the determination of COVID-related fatalities. However, it's important to note that not all COVID-related deaths may be accurately identified or reported, leading to potential underestimation or misclassification in official statistics.

Incorporating this understanding into our analysis, our measurement for COVID death rates is drawn from a accumulated dataset last updated until March 2023. This temporal constraint means we would not be aware of the fluctuations of death rates over time, particularly during the peak of the pandemic in 2020. The data's primarily reliance on the reported statistics from hospitals, clinics, testing centres, and laboratories may introduces bias as excludes uncounted deaths from COVID. Especially during the critical period of health facilities capacity constraints,many individuals died outside of medical facilities without being documented. Therefore, our measurements for COVID death rates are only accounted for the documented death from COVID.

In addition,it's important to note that the death rate presented here reflects the entire population of a state. However, COVID-19 disproportionately affects older individuals compared to younger adults in the prime of their lives. As a result, this data does not capture the effect of COVID across different age groups. Instead, it only offers a general overview the state's population. In summary, we are unable to provide detailed insights into age-specific or temporal variations based on this dataset.

Similarly, our measurement of life expectancy suffers from demographic limitations. Life expectancy, defined as the estimate of the average age that members of a particular population group will be when they die, is a crucial metric in understanding population health dynamics [@citedefenition]. Even though we categorize life expectancy by race, the absence of an age breakdown restricts our ability to discern impacts across different age groups. Moreover, delays in reporting deaths undermine the timeliness and accuracy of our data, potentially skewing our understanding of population health trends.

Our measurement for each state's political preference is calculated from the 2020 election votes. There may be many scenarios that prevented eligible voters from voting. However, we are only able to assess the the state's political preference based on the people who voted, thereby potentially introducing selection bias. Additionally, issues such as ballot loss or voter disenfranchisement further compound the reliability of our measurements.

### Measurement Techniques

Manual Data Extraction involved careful review of reports and tables to ensure accurate data capture, particularly for datasets like life expectancy where this meticulous approach was paramount; this was coupled with Standardization and Adjustment processes where excess death data and age-standardized excess death rates underwent significant processing, including the comparison of observed deaths during the pandemic to expected deaths based on historical trends, with adjustments made for the age distribution of the population; concurrently, Geospatial Data Handling was employed for mapping the GHS Index scores, utilizing Geographic Information Systems (GIS) to visually represent pandemic preparedness across various countries, and this was complemented by Real-Time Data Aggregation, which entailed aggregating COVID-19 death data in real-time from authoritative sources such as the Johns Hopkins University Dashboard, necessitating continuous data monitoring and updates to maintain accuracy and relevance.

### Adjustments and Considerations

Accounting for underreporting through the measurement of excess deaths provides a more comprehensive view of the pandemic's impact by capturing deaths that may have been missed or misclassified as COVID-19, while the process of age standardization adjusts these excess death rates to a standard age distribution, enabling meaningful comparisons between countries with varying demographic profiles. However, the analysis had to navigate challenges associated with varying data quality and availability across different countries, especially given the diverse COVID-19 reporting standards and testing capacities. Moreover, the study also acknowledges the significant influence of political and social factors on the pandemic's outcomes, which, despite being more challenging to quantify, are essential for a holistic understanding of the pandemic's impact.

# Results {#sec-results}

Our results are summarized in the following figures. @fig-ELE illustrates the trend of life expectancy at birth across different racial groups over time. Unfortunately, due to the unavailability of data over time for Asian, American Indian, and Alaska Native communities, they have been excluded from the time series graph. @tbl-life-exp provides a more detailed breakdown of the life expectancy before and during COVID for Asian, American Indian, Alaska Native and other racial communities.

An intriguing observation from @fig-ELE is the consistently higher life expectancy among Hispanic individuals compared to other groups, even amidst the challenges posed by COVID-19. On the other hand, Black individuals have consistently exhibited lower life expectancy, which further declined notably in 2021, reaching just over 70.8 years old. The life expectancy trends of white people and all other races and origins remain close together throughout the 14 years, with minimum variance.

While all racial groups experienced declines in their average life expectancy, the declines vary greatly. The life expectancy for White individuals decreased from 78.8 to 76.4 years old from pre-pandemic levels in 2019 to 2021, while for Hispanic individuals, it dropped by 4.2 years, and for Black individuals, it declined by 4 years during the same period. Our data aligns with the findings of @andrasfay_goldman_2022, which states that the Black and Hispanic populations experienced twice as much life expectancy decrease as the white population."

```{r fig.width=7, fig.height=3.5}
#| echo: false
#| eval: true
#| label: fig-ELE
#| fig-cap: Estimates of Life Expectancy at Birth, by Race 2006-2021
#| warning: false

# Reshape the data if necessary
life_exp_melted <- reshape(
  life_exp,
  idvar = "year_id",
  varying = list(names(life_exp)[-1]), 
  v.names = "life_exp",
  times = names(life_exp)[-1],
  new.row.names = 1:1e5,
  direction = "long"
)
life_exp_melted$race <- gsub("_", " ", life_exp_melted$time)  # Convert "_" to " " in race names
life_exp_melted <- life_exp_melted[, c("year_id", "race", "life_exp")]

life_exp_melted[life_exp_melted$race == 'All', "race"] <- 'All races and origins'

## CREATE PANEL A
est_life <- ggplot(data = life_exp_melted[!(life_exp_melted$race %like% 'AIAN|Asian'), ], aes(x = year_id, y = life_exp, color = race)) +
  geom_point(shape = 16, size = 3.5, alpha = 0.9) +  # Circular points with size 2
  geom_line(aes(group = race), alpha = 0.56, size = 1.5) +  # Add group aesthetic for lines
  scale_y_continuous(name = 'Life expectancy at birth', limits = c(65, 85)) +
  scale_x_continuous(name = '', breaks = seq(2006, 2020, by = 2), limits = c(2005.5, 2021.5)) +
  scale_color_manual(values = c('#0c4928', '#0b3669', '#4c2962', '#5a0f1c'), 
                     labels = c("All races and origins", "Hispanic", "Black", "White")) +
  # Darker color values
  theme_minimal() +
  theme(axis.title = element_text(size = 14),  # Adjusting the size of the y-axis label
        axis.text = element_text(size = 12),
        legend.position = 'bottom',  # Set legend position to bottom
        legend.justification = 'center',  # Center the legend
        legend.text = element_text(size = 10),  # Adjusting the font size for legend text
        legend.title = element_text(size = 12, face = 'bold')) +
  labs(color = "Race:")  # Set the legend title

est_life

```

```{r}
#| echo: false
#| eval: true
#| warning: false
# Create a dataframe to store the changes by race
change_by_race <- data.frame(
  race = c("All races and origins", "Hispanic", "American Indians and Alaska Native", "Asian", "Black", "White"),
  change_2020_2019 = c(77 - 78.8, 77.9 - 81.9, 67.1 - 71.8, 83.6 - 85.6, 71.5 - 74.8, 77.4 - 78.8),
  change_2021_2020 = c(76.1 - 77, 77.7 - 77.9, 65.2 - 67.1, 83.5 - 83.6, 70.8 - 71.5, 76.4 - 77.4)
)

```

```{r fig.width=8, fig.height=4}
#| echo: false
#| eval: true
#| label: fig-CLE
#| fig-cap: Change in Life Expectancy at Birth from the Previous Year
#| warning: false

# Melt the dataframe
tmp <- melt(change_by_race, id.vars = "race", measure.vars = c("change_2020_2019", "change_2021_2020"), 
             variable.name = "period", value.name = "diff")

# Create PANEL B
change_life_exp <- ggplot(data = tmp, aes(x = race, y = diff, fill = period)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.71) +
  scale_y_continuous(name = "Change (years)", breaks = -5:0, limits = c(-5.5, 0)) +
  scale_fill_manual(name = "Time period", values = c('#243f29', '#186d38'), labels = c('2019-2020', '2020-2021')) + 
  scale_color_manual(name = "Time period", values = c('#243f29', '#186d38'), labels = c('2019-2020', '2020-2021')) + 
  theme_minimal() + labs(x = '') +
  geom_text(aes(label = sprintf("%.1f", diff)), position = position_dodge(width = 0.9), vjust = 1.3, size = 4) +
  theme(
    axis.title = element_text(size = 14),
    axis.text.x = element_text(size = 10, hjust = 1),
    axis.text.y = element_text(size = 8),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12, face = 'bold')) + 
    scale_x_discrete(labels = function(x) str_wrap(x, width = 15))

change_life_exp

```

```{r}
#| echo: false
#| eval: true
#| warning: false

vote_data_by_state <- vote_share %>%
  group_by(state_name) %>%
  summarize(
    total_votes = sum(total_votes),
    total_gop_votes = sum(votes_gop),
    total_dem_votes = sum(votes_dem),
    total_diff = sum(diff),
    avg_per_gop = mean(per_gop),
    avg_per_dem = mean(per_dem),
    avg_per_point_diff = mean(per_point_diff)
  )

vote_data_by_state <- vote_data_by_state |>
  mutate(gop_greater = if_else(total_gop_votes > total_dem_votes, 1, 0))

```

```{r}
#| echo: false
#| eval: true
#| label: fig-Vote
#| fig-cap: Total Number of Votes in Top 10 States with Highest Proportion of Republican Votes 
#| warning: false

# Calculate proportion of Republican votes
sorted_states <- vote_data_by_state
sorted_states$proportion_republican <- sorted_states$total_gop_votes / sorted_states$total_votes

# Sort by proportion of Republican votes
sorted_states <- sorted_states[order(-sorted_states$proportion_republican), ]

# Select top 10 states
top_10_states <- head(sorted_states, 10)

# Extract the top 10 states with the biggest difference
top_10_states <- head(sorted_states, 10)
# Create a data frame for plotting
plot_data <- data.frame(
  state_name = top_10_states$state_name,
  republican_votes = top_10_states$total_gop_votes,
  democrat_votes = top_10_states$total_dem_votes
)

# Convert data to long format for ggplot
plot_data_long <- tidyr::pivot_longer(plot_data, cols = c(republican_votes, democrat_votes), names_to = "party", values_to = "votes")

# Create the stacked bar plot
ggplot(plot_data_long, aes(x = state_name, y = votes, fill = party)) +
  geom_bar(stat = "identity") +
  labs(x = "State", y = "Total Votes", fill = "Party") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  scale_fill_manual(values = c("republican_votes" = "red", "democrat_votes" = "blue"),
                    labels = c("Democratic Votes", "Republican Votes"))

```

```{r}
#| echo: false
#| eval: true
#| label: tbl-deaths
#| tbl-cap: Top 10 States with Highest Death Rates from COVID-19 (per 100,000 people)
#| warning: false

# Assign column names
colnames(deaths) <- c("State", "Death_Rate")

# Rank the states based on death rate
deaths <- deaths[order(-deaths$Death_Rate), ]
deaths$Rank <- 1:nrow(deaths)

# Create the table
table <- head(data.frame(Rank = deaths$Rank, State = deaths$State, Deaths = deaths$Death_Rate), 10)

# Create a gt table
table_gt <- gt(table)

# Print the gt table
table_gt

```

```{r}
#| echo: false
#| eval: true
#| label: tbl-life-exp
#| tbl-cap: Life Expectancy by Race (2019-2021)
#| warning: false

# Filter the data for the years 2019 to 2021
life_exp_subset <- life_exp_melted[life_exp_melted$year_id >= 2019 & life_exp_melted$year_id <= 2021, ]
# Pivot the data to create a table and rename columns
life_exp_table <- reshape2::dcast(life_exp_subset, year_id ~ race, value.var = "life_exp")
colnames(life_exp_table) <- c("Year", "All races and origins", "Hispanic", "AIAN", "Asian", "Black", "White")
# Create a gt table
life_exp_gt <- gt(life_exp_table, rowname_col = "Race") 
# Print the gt table
life_exp_gt

```

@fig-CLE serves as a valuable addition to our previous analysis with the inclusion of Asians, American Indians, and Alaska Natives during the critical period from 2019 to 2021. With these additional ethnic categories, we can discern that American Indians and Alaska Natives experienced significant impacts from the pandemic, with a decrease of 4.7 years in the first year and 1.9 years in the second year, totaling 6.6 years.

With a chart that depicts the change in life expectancy each year by ethnicity, we can better gain a more comprehensive understanding of the impact of COVID by ethnicity. While the declines in life expectancy are smaller in magnitude from 2020 to 2021, they are notably minimal for Asians and Hispanics, with reductions of -0.1 and -0.2 years respectively.

Various studies have confirmed the existence of correlation between a state's political affiliation and its handling of COVID issues (@datz_2022). To testify to the claim, we collected the voting data for each of the 50 states for the 2020 presidential election. We then selected the ten states with the highest proportions of Republican votes. These voting patterns are visualized in @fig-Vote, where red symbols represent Republican votes and blue symbols represent Democratic votes. Among the top ten states, Tennessee recorded the highest total number of votes, while Wyoming boasted the highest proportion of Republican votes.

Given the variation in population sizes among states, direct comparisons of COVID-19 case numbers are inherently flawed. As an alternative, we employed death rates per 100,000 people as a metric for evaluating each state's COVID preparedness and situation. We then produced @tbl-deaths that rank the top ten states with the highest death rates from COVID-19 per 100,000 people to examine the potential correlation between party preferences and COVID-related deaths. Notably, We found that six of the ten top Republican states made a reappearance in the top death rates table; these states are Oklahoma, West Virginia, Arkansas, Alabama, Tennessee, and Kentucky. The higher mortality rates observed in Republican-leaning states may be attributed to their preference for less stringent measures compared to Democratic-leaning states, which tend to favor stricter measures (@vandusky-allen_shvetsova_2021).

Following our previous analysis regarding individuals' political affiliations, we have developed @fig-deaths-rates-votes, which encompasses all 50 states of the US along with their political leanings based on which party garnered the majority votes. This information is juxtaposed against their respective COVID-19 death rates.

While we cannot make any definitive assertions about stark differences, we do observe that the Republican-leaning states are slightly more clustered around higher death rates ranging from  350 to 450 deaths, while the Democratic-leaning states appear to be more evenly distributed, and notably one Democratic-leaning state has the lowest death rate. An intriguing observation is that although many Republican-leaning states demonstrate higher COVID death rates, it is noteworthy that Arizona, typically considered a Democratic-leaning state, records the highest death rate among all states.

```{r}
#| echo: false
#| eval: true
#| label: fig-deaths-rates-votes
#| fig-cap: COVID-19 Death Rates vs. Republican Votes
#| warning: false
merged_data <- merge(deaths, vote_data_by_state, by.x = "State", by.y = "state_name")
# Create a scatter plot of COVID-19 death rates against the percentage of Republican votes
ggplot(merged_data, aes(x = as.factor(gop_greater), y = Death_Rate)) +
  geom_point(shape = 21, fill = "skyblue", color = "darkblue", size = 3) + # Use filled circles with blue color
  labs(x = "Republican Votes", y = "COVID-19 Death Rate") +
  theme_minimal() + # Use minimal theme for cleaner appearance
  theme(plot.title = element_text(size = 16, face = "bold"), # Adjust title appearance
        axis.title = element_text(size = 14)) + # Adjust axis labels appearance
  scale_x_discrete(labels = c("Democratic Majority", "Republican Majority")) # Set custom labels for x-axis
```

# Discussion {#sec-discussion}

This begs the question as to why we are seeing these results. There isn't exactly a single answer to this question, however, we can certainly point out some considerable factors to this result. These factors can be broken down into the influence of political polarization, the impact of government transparency which ties into the effect it also had on public trust, and the role of social vulnerabilities and healthcare access disparities. We should also look at how to address these potential factors by implementing strategies in the future in regards to data collection and sharing. It is important to also note we were able to identify some weaknesses in our research that should be evaluated in the future.

## Influence of political polarization on adherence to health guidelines

Political polarization has significantly impacted the adherence to health guidelines during the COVID-19 pandemic. The divergence in political ideologies has translated into differing attitudes towards health directives, including mask mandates, social distancing, and vaccination uptake.

Various studies and our results have consistently demonstrated a correlation between political affiliation and adherence to health guidelines during the COVID-19 pandemic. This correlation extends to distinct behaviors and compliance levels with health recommendations, directly impacting COVID-19 case rates and mortality. One study, in particular, revealed a noteworthy trend: while Democratic-leaning states initially faced higher cases and death rates at the onset of the pandemic, Republican-leaning states ultimately suffered more severely throughout the latter half of 2020 [@march_11_2021_2021]. This shift in severity can be attributed to several factors. Initially, many Democratic states were the epicenter of the virus, experiencing high case and death rates due to the initial entry of the virus into these regions. However, as the pandemic progressed, Republican-leaning states saw a surge in cases, partially due to their relatively loose COVID-19 regulations, such as the lack of strict mask requirements [@march_11_2021_2021].

The adherence to health guidelines is evident in the varied health outcomes observed across the United States. Regions with lower compliance to health directives, often influenced by political leanings, have experienced higher rates of COVID-19 transmission, hospitalizations, and deaths. The disparities in vaccine uptake, driven by political affiliations, have further exacerbated these outcomes, leaving certain communities more vulnerable to the virus and its variants. To mitigate the influence of political polarization on public health, it is imperative to depoliticize health guidelines and focus on evidence-based approaches to disease prevention and control. Building trust in health institutions and promoting bipartisan support for public health measures are essential steps toward achieving higher compliance and better health outcomes. Engaging trusted community leaders and utilizing targeted communication strategies can also help bridge the divide and encourage adherence to health guidelines.

## Impact of government transparency and consistent communication on public trust.

The politicization of health guidelines and mixed messages from political and health leaders during the COVID-19 pandemic have significantly undermined the effectiveness of public health messaging, leading to confusion, skepticism, and eroded trust among the public. Initially, inconsistencies in recommendations, such as on mask usage, challenged the principle of clear, consistent, and science-based communication essential for an effective public health response.

Moreover, the transparency of government actions and decision-making processes is crucial in building and maintaining public trust, especially during health crises. The level of public trust was greatly affected by the openness and accuracy with which governments, at all levels, communicated about the evolving situation, the reasoning behind guidelines, and the measures taken to combat the virus, emphasizing the importance of transparent reporting of data related to case counts, hospitalizations, vaccine distribution, and side effects.

Furthermore, consistent communication from public health officials and government leaders is key to ensuring adherence to health guidelines, where inconsistencies, such as changes in mask-wearing guidelines without clear explanations, have led to public confusion. The direct correlation between government transparency, consistent communication, and public behavior is self-evident, with populations receiving clear and transparent information being more likely to adhere to guidelines, participate in testing and tracing efforts, and accept vaccination.

Drawing lessons from the pandemic, strategies for improving government transparency and communication in future health emergencies should include establishing centralized information hubs, ensuring regular and predictable communication from health authorities, engaging community leaders in information dissemination, and harnessing digital platforms and social media to amplify public health messages, thus reinforcing public trust and compliance.

## Role of social vulnerabilities and healthcare access disparities in pandemic impact

The COVID-19 pandemic starkly highlighted how social vulnerabilities and disparities in healthcare access exacerbated the impact of global health crises, contributing to significant variations in disease outcomes and underscoring the need for targeted public health strategies that address these disparities' root causes.

Social vulnerabilities, such as socioeconomic status, race, ethnicity, and housing conditions, critically determined COVID-19 outcomes' severity, with populations in crowded housing, limited access to sanitation, and lower socioeconomic brackets experiencing higher transmission rates due to social distancing and hygiene maintenance challenges. An independent study done by the `Government of Canada` states that it "identified that the risk of COVID-19 related deaths in Black, Asian and minority ethnic groups was nearly 1.5 times higher than White individuals" [@citesource5]. Another news article done by the `MSN` [@citesource6] states that 'racial minority participants reported more negative impacts on health status, activity, and absence from work as compared to the White population.'

The pandemic's economic toll further limited these groups' healthcare access, amplifying vulnerabilities. Disparities in healthcare access played a significant role in influencing COVID-19 morbidity and mortality, with communities facing healthcare facility shortages, provider scarcities, and barriers due to insurance or financial constraints at heightened risk. These disparities were evident in the uneven vaccine distribution and access, highlighting the advantages of regions with strong healthcare infrastructure.

Marginalized populations, including racial and ethnic minorities, faced compounded risks from social vulnerabilities and healthcare disparities, evidenced by higher infection, hospitalization, and death rates due to factors like essential service employment and prevalent pre-existing conditions. A news article from `CNN` [@citesource7] talks about how many essential workers, who cannot work from home, are from black and Latinx communities. These include healthcare professionals, grocery cashiers, delivery workers, and public transport employees. Despite their crucial roles, they often lack adequate pay, protection, and respect.

Addressing these disparities in future pandemics requires public health strategies that prioritize equity and inclusivity, including community-based healthcare investments, enhanced vulnerable community outreach, and equitable healthcare resource access policies. By incorporating social determinants of health into public health preparedness plans, responses can effectively protect at-risk populations, making future public health responses more resilient, inclusive, and effective in safeguarding all population segments.

## Strategies for improving real-time data collection and sharing for public health decisions.

To address the fragmentation in data collection and sharing witnessed during the pandemic, it's crucial to establish integrated data platforms that enable seamless health data exchange among various health agencies and stakeholders, utilizing cloud computing and APIs for real-time accessibility and usability. Equally important is enhancing data standardization and interoperability through universal standards like FHIR (Fast Healthcare Interoperability Resources) to facilitate efficient data sharing [@citesource8].

Another notable example is the Digital Health Platform (DHP) [@citedhp] developed by the European Centre for Disease Prevention and Control (ECDC) [@citeecdc], which facilitates the real-time exchange and analysis of health data across EU member states. This platform has been instrumental in coordinating responses to health threats, including the COVID-19 pandemic, by providing timely and accurate data to inform public health decisions.

Another digital platform worth mentioning is the HealthMap system [@citehealthmap], which utilizes online data sources and machine learning algorithms to detect and track global disease outbreaks. HealthMap's ability to aggregate and visualize data from various sources, including news reports and social media, has proven invaluable in early outbreak detection and response, showcasing the power of digital surveillance in public health.

Investing in digital surveillance systems, which employ AI and machine learning to sift through diverse data sources for early outbreak detection, is essential for rapid response to health threats. An news article by `nature` [@citegoogle] talks about how Google has made an AI which claims to be "better bedside manner than human doctors — and makes better diagnoses". Furthermore, fostering public-private partnerships can harness the agility of the private sector and the public health expertise of governmental agencies to enhance data analytics capabilities.

Ensuring the privacy and security of health data through robust governance frameworks and advanced encryption is paramount to maintaining public trust. Engaging communities in these initiatives ensures their relevance and fosters trust while building global data-sharing networks encourages international collaboration, crucial for a concerted response to pandemics.

Collectively, these strategies are fundamental to bolstering public health decision-making and preparedness, making our health systems more resilient against the challenges posed by emerging infectious diseases.

## Weaknesses and next steps

In our research, we identified several weaknesses in our data that is worth mentioning. The pandemic highlighted significant disparities in global surveillance, with countries possessing advanced healthcare systems paradoxically reporting higher COVID-19 mortality rates due to more accurate detection capabilities, underscoring the need for universal surveillance standards to accurately monitor diseases worldwide.

The United States, in particular, faced challenges with delayed data reporting, impeding swift public health responses and decision-making. The effects of this is self-evident as it makes us question if the data we are given is even correct. Furthermore, the often incomplete demographic data obscured the unequal effects of the pandemic on diverse communities, preventing targeted interventions for those most at risk, such as racial and ethnic minorities, thereby deepening existing health inequities.

Additionally, the pandemic revealed a critical lack of infrastructure capable of integrating various data types, from clinical to public health surveillance, limiting a comprehensive understanding of the pandemic's scope and the impact of measures taken. Addressing these data weaknesses necessitates establishing global data collection, reporting, and sharing standards, including case definitions, demographic information, and data formats to ensure consistency across borders.

Investments in digital health technologies are essential for enabling real-time surveillance while maintaining privacy and security. Policies should be enacted to ensure the collection of detailed demographic data to identify and address health disparities, ensuring equitable intervention distribution. Developing frameworks that allow for the seamless integration of diverse data sources will offer a more detailed pandemic perspective, necessitating cross-sector collaboration.

Enhancing data literacy among health professionals, policymakers, and the public will foster informed decision-making and trust in health interventions. Lastly, international collaboration in data sharing and analysis is crucial for the rapid dissemination of insights and best practices, bolstering global preparedness and response for future health crises.

# Conclusion {#sec-conclusion}

Despite the United States' top ranking in the Global Health Security Index as the most prepared nation for pandemics, its actual response to the COVID-19 pandemic fell short of expectations, leading to disproportionately high death rates. This discrepancy is attributed to several critical factors highlighted in the paper.

The U.S. did not fully capitalize on its pandemic preparedness resources, with early testing blunders and a lack of a unified national testing approach hindering its response. Additionally, inherent vulnerabilities such as the significant portion of the population in congregated settings like nursing homes and prisons increased susceptibility to virus spread and severe health outcomes.

The politicization of the pandemic response further compounded these issues, resulting in varied adherence to public health guidelines across states and political lines, thereby undermining the response's effectiveness. Moreover, inconsistent public health communications and the challenge of accessing standardized, quality data impeded the implementation of localized, effective interventions. The pandemic also highlighted and intensified existing socio-economic disparities, disproportionately impacting marginalized communities who faced higher risks and adverse outcomes, a situation that the response efforts failed to sufficiently mitigate.

\newpage

# References